---
title: "demo"
author: "Gabriel Krotkov"
format: html
---

# Getting Started

Hello! I'm Gabriel, and I'll be using this rmarkdown to give you a demo of the useful functions that scoutR provides to use TBA/Statbotics data in R. 

For those using a QMD file for the first time, QMD stands for "Quarto Markdown", it is a file type that blends formattable text, chunks of code, and output. To run the code while viewing a qmd, click the green arrow on the top right corner of a code chunk.

## Installation

If you do not already have `devtools` installed, you will need to install that first. The below line of code installs devtools. If you're new to R programming, I also recommend installing the package `tidyverse`; it is an extremely large package and takes a long time to install, but having access to the Tidyverse will make your R programming life much easier.

Installation code should only be run once. Once you have run the installation on your machine, I suggest commenting out installation code so you don't accidentally rerun that code by running the entire file.

```{r}
# only run this code once!
install.packages("devtools")
```

To install `scoutR`, run the following line of code in your R console.

```{r}
devtools::install_github("gkrotkov/scoutR")
```

## Initialization

Next, you will need to initialize scoutR with your TBA auth key. You can generate a TBA authorization key at www.thebluealliance.com/account under the "Read API Keys" header. Once you have an auth key, replace the string in this code with your auth key! (The auth key needs quotations around it)

```{r}
library(scoutR)
initialize_scoutR("<replace this with your auth key!>")
```

# Functions Demo

scoutR provides a *lot* of handy functions. Here, I'll show you some of the ones I find myself using the most. 

@TODO do a tour of key functions

- event_matches()
- get_multifield_df()
- 

## get_multifield_df()

I need to come up with a better name for this function; but it keeps being handy so I want to highlight it. `get_multifield_df()` returns a dataframe 

```{r}
vagle <- get_multifield_df(event_matches("2025vagle"))
```

