---
title: "tbaR demo"
author: "Gabriel Krotkov"
date: "2023-04-28"
output: 
  html_document:
    code_folding: hide
---

# Introduction

Hello! I'm Gabriel, and I'll be using this rmarkdown to give you a demo of the useful functions that tbaR provides for anyone interested in interacting with TBA data in a convenient format. 

If this is your first time using tbaR, you may need to install some packages. You can install a package by running code like below. You should only run the installation once, but loading a library with the `library` function will be run many times. (The tidyverse is a large package, don't worry if it's taking a long time to install.)

```{r, eval = FALSE}
install.packages(c("devtools", "tidyverse", "stringr", "jsonlite", "assertthat" 
                   "httr", "sys", "stringr", "roxygen2", "rvest", "data.table"))
```

The below chunk of code loads all of tbaR's dependencies (implicitly) as well as all of tbaR's files. It also executes the document() function while creates the documentation of the tbaR files.

```{r, echo = FALSE}
my_auth_key <- read_file("auth_key.txt")
```


```{r}
library(devtools)
load_all()
initialize_tbaR(my_auth_key)
```

If you wanted to utilize tbaR from outside the tbaR directory, you would use code in this form.

```{r, eval = FALSE}
# for example, this code would work if called from a file that sees the directory "tbaR"
path_to_tbaR <- "tbaR"
library(devtools)
load_all(path_to_tbaR)
```

# tba_interfaceR

InterfaceR (pronounce "interfacer") is the primary product of tbaR. It provides R functions that connect to the TBA API. Each function returns a tidied dataframe of data instead of the JSON output of the TBA API.

```{r}
awards_example <- team_awards(321, year = 2023)


```

@TODO multievent_insights in interfaceR, some solution to the developer key problem

# tba_helpR

Helper (pronounce "helper")

```{r}

```

# cOPR plotting

This code chunk is an example of directly scraping the TBA website to get cOPR data, which is available on the website's frontend but not through the API (at least yet.)

```{r}
event_key <- "2023mil"
df <- event_coprs(event_key)

df %>% mutate(cargo_ratio =`Total Cones Scored`-`Total Cubes Scored`) %>% 
    ggplot(aes(label=team,
               x=`Total Cones Scored`, 
               y=`Total Cubes Scored`,fill=cargo_ratio,
               color=)) + 
    ggrepel::geom_label_repel() + 
    scale_fill_gradient2(high = 'yellow', mid = 'white', low = 'purple') +
    geom_abline(slope = 1) +
    ggtitle("COPR cone vs cube ratio")

#or the averages computed by tba from website
raw <- GET(paste0("https://www.thebluealliance.com/event/", 
                  event_key, "#results")) %>%
    xml2::read_html() %>% html_table()

df2 <- raw[[24]]
df2 %>% select(Team, 5) %>% ggplot(aes(x=reorder(factor(Team),`Avg Charge Station`),y=`Avg Charge Station`))+ geom_col() + coord_flip() +labs(y='team',caption='source: TBA 2023txfor website https://www.thebluealliance.com/event/2023txfor#results')
```
