---
title: "Intro to scoutR"
author: "Gabriel Krotkov"
date: "2025-04-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
library(scoutR)
```

# Getting Started

Hello! This is your introduction to using `scoutR`, assuming you might even be a first-time R user. `scoutR` is a library for the R programming language designed to facilitate scouting & data science for the FIRST Robotics Competition.

## R and RStudio

`scoutR` is a package for the R programming language, an open-source statistical programming language popular for its support for statistical programming, data science, and data visualization. The most typical IDE used for R programming is RStudio, published by Posit. You can download and install the R programming language and RStudio [here](https://posit.co/download/rstudio-desktop/). You can use other IDEs (Visual Studio Code is a popular one, for example) to program in R, but RStudio is specifically designed to support R programming.

## Package Installation

If you do not already have `devtools` installed, you will need to install that first. To do that, run this line of code in your R console:  

```{r, eval = FALSE}
# only run this code once!
install.packages("devtools")
```

Installation code should only be run once. Once you have run the installation on your machine, I suggest deleting or commenting out installation code so you don't accidentally rerun that code by running the entire file.

Once you have installed `devtools`, you can install `scoutR`. To install `scoutR`, run the following line of code in your R console.

```{r, eval = FALSE}
devtools::install_github("gkrotkov/scoutR")
```

## Initialization

Next, you will need to initialize scoutR with your TBA auth key. You can generate a TBA authorization key at www.thebluealliance.com/account under the "Read API Keys" header. Once you have an auth key, replace the string in this code with your auth key! (The auth key needs quotations around it)

```{r, eval = FALSE}
library(scoutR)
initialize_scoutR("<replace this with your auth key!>")
```

# Functions Demo

Once you've installed and initialized `scoutR`, you can start using some of the handy functions it provides! Here are some of the functions I find myself using the most at competition. For every exported function in `scoutR`, you can run `help(fxn_name)` to pull up documentation listing the function's usage, parameters, and some examples.

## event_matches()

I find this function to be the real workhorse for scouting applications. This function takes a TBA-legal event key (examples: "2023mil", "2025njtab") and returns a dataframe with data for the matches played at that event. A key parameter for this function is `match_type`, which can be "all", "qual", or "playoff".

```{r}
mil23 <- event_matches("2023mil")
# use match_type to subset down to only the matches you want
mil23_quals <- event_matches("2023mil", match_type = "qual")
mil23_playoffs <- event_matches("2023mil", match_type = "playoff")

# take a look at all the useful data!
head(mil23)
```

## event_tangibles()

`event_tangibles()` allows you to pull all the single-robot records from TBA in one line of code, based on the standard TBA naming schema since 2018. 

```{r}
# defaults to only pulling data from qualification matches
vagle25_tangibles <- event_tangibles("2025vagle")
# but if you suppress the qual_only flag, you'll get playoffs as well
vagle25_tangibles <- event_tangibles("2025vagle", qual_only = FALSE)
head(vagle25_tangibles)
```

## events()

Retrieve all the events in a given year, with flags to get key-only return or official-only events

```{r}
chargedup_events <- events(2023)
head(chargedup_events)
chargedup_official_events <- events(2023, official = TRUE)
head(chargedup_official_events)
crescendo_keys <- events(2024, keys = TRUE)
head(crescendo_keys)
```

## qual_schedule()

Retrieve the posted qual schedule. Usually, you'll want to direct this to a CSV for use in whatever scouting application you're using that depends on the schedule. 

```{r}
mdpas <- qual_schedule("2025mdpas")
head(mdpas)
# write.csv(mdpas, file = "2025mdpas_schedule.csv")
```

## team_awards()

Get all the awards won by a given team. Includes optional parameters to filter for year or event.

```{r}
gos_awards_history <- team_awards(3504)
head(gos_awards_history)
```

## event_coprs()

Retrieve all component OPRs posted by TBA on the "Insights" page. 

```{r}
vagle25_coprs <- event_coprs("2025vagle")
head(vagle25_coprs)
```

## event_season_history()

This function is useful mostly for prescouting - it retrieves robot-level results for every team registered for a given event. Note that this differs from `event_tangibles` in that this function gets *all* the matches a team has played this season, while `event_tangibles` focused on what happened *at a particular event.*

```{r}
newton24_prescout <- event_season_history("2024new")
head(newton24_prescout)
```
