---
title: "OPR GLM"
format: html
---

Key Question: Can we improve OPR's fit to the data by using a GLM? 

I would expect 2018 to be the best candidate year for this method, so we'll test using that at the beginning to see if we can get a meaningful enough improvement to continue investigating. 

```{r}
rm(list = ls())
library(devtools)
load_all()

gpr <- event_matches("2024paca", match_type = "qual")
load("../data/opr_binvar.rda")

linear_fit <- fit_lineup_lm(
    gpr, 
    list(red = gpr$red_score, blue = gpr$blue_score), 
    w_bins
)

plot(linear_fit, which = 1:6)
```

```{r}
design <- lineup_design_matrix(gpr)
response <- c(gpr$blue_score, gpr$red_score)
rlm_fit <- rlm(design, response)
```

# Regularization

```{r}
rm(list = ls())
library(devtools)
library(glmnet)
load_all()

match_df <- event_matches("2023mil", match_type = "qual")
design <- data.matrix(lineup_design_matrix(match_df))
# could adjust this response to win margin; would be answering a different
# question.
response <- c(match_df$blue_score, match_df$red_score)

# grid of lambdas to consider
lambda_seq <- 10 ^ seq(-2, 2, by = .05)

# Using cross validation glmnet
ridge_cv <- cv.glmnet(design, response, alpha = 0, lambda = lambda_seq)
# refit the ridge using the min-MSE lambda
best_ridge <- glmnet(design, response, alpha = 0, lambda = ridge_cv$lambda.min)

# Try a LASSO - could LASSO be used to kinda "auto-tier" teams?
lasso_cv <- cv.glmnet(design, response, alpha = 1, lambda = lambda_seq)
best_lasso <- glmnet(design, response, alpha = 1, lambda = lasso_cv$lambda.min)

# refitting ridge & LASSO with the min-MSE lambda but insisting on regression 
# through the origin
origin_ridge <- glmnet(
    design, response, alpa = 0, lambda = ridge_cv$lambda.min, intercept = FALSE
)

origin_lasso <- glmnet(
    design, response, alpha = 1, lambda = lasso_cv$lambda.min, intercept = FALSE
)

results <- cbind(
    coef(best_ridge), coef(best_lasso), coef(origin_ridge), coef(origin_lasso)
)
```

Notably - without a constant, LASSO is unable to push many coefficients to exactly 0! It basically becomes very similar to ridge regression.

# Elastic Net Regression

```{r}

```

