---
title: "Playground"
format: html
---


```{r}
# loading the scoutR library: 
# https://github.com/GKrotkov/scoutR
library(devtools); load_all()
```

# WTF is up with 2022iscmp, 2022chcmp, and 2022micmp3?

```{r}
rm(list = ls())
iscmp <- event_matches("2022iscmp", match_type = "qual")
chcmp <- event_matches("2022chcmp", match_type = "qual")
aptiv <- event_matches("2022micmp3", match_type = "qual")
```

Looking at this event report, none of these match what `TBA Insights` shows (they are pretty close, usually within 0.5 - however the problems are more pronounced for APTIV.)

```{r}
event_report <- function(evnt, label){
    margin <- abs(evnt$red_score - evnt$blue_score)
    winning_score <- pmax(evnt$red_score, evnt$blue_score)
    scores <- c(evnt$red_score, evnt$blue_score)
    cat("For the", label, "event, the data shows:\n", 
        "Average score -", mean(scores), "\n", 
        "Average winning score -", mean(winning_score), "\n", 
        "Average winning margin -", mean(margin), "\n")
}

event_report(iscmp, "Israel DCMP")
event_report(chcmp, "CHS DCMP")
event_report(aptiv, "Aptiv Division")
```

Let's take a look at APTIV specifically since the disagreement is the most pronounced there. 

```{r}
margin <- abs(aptiv$red_score - aptiv$blue_score)
winning_score <- pmax(aptiv$red_score, aptiv$blue_score)
```

Manual inspection of the scores doesn't reveal a difference between the scores listed on TBA and the scores output by the API.

```{r}
library(rvest)

report_from_tbl <- function(tbl){
    red_scores <- tbl %>%
        html_elements(".redScore span") %>%
        html_text() %>%
        as.numeric()
    red_scores <- red_scores[seq(1, length(red_scores), by = 2)]
    
    blue_scores <- tbl %>%
        html_elements(".blueScore span") %>%
        html_text() %>%
        as.numeric()
    blue_scores <- blue_scores[seq(1, length(blue_scores), by = 2)]

    cat("Avg:", mean(c(red_scores, blue_scores)), "\n", 
        "Winning Score Avg:", mean(pmax(red_scores, blue_scores)), "\n",
        "Margin:", mean(abs(red_scores - blue_scores)), "\n")
}

discrepancy_report <- function(url){
    html <- read_html(url)
    
    cat("\nQUALS:\n")
    report_from_tbl(html_elements(html, "#qual-match-table"))

    cat("\nPLAYOFFS\n")
    report_from_tbl(html_elements(html, "#elim-match-table"))
}

discrepancy_report("https://www.thebluealliance.com/event/2022chcmp#results")
discrepancy_report("https://www.thebluealliance.com/event/2022iscmp#results")
discrepancy_report("https://www.thebluealliance.com/event/2022micmp3#results")

```

# Double Harmonies?

```{r}
cresc_evnts <- events(2024, keys = TRUE)
cresc_matches <- lapply(cresc_evnts, event_matches)
cresc_matches <- cresc_matches[!is.na(cresc_matches)]

get_harmonies <- function(df){
    return(c(df$blue_endGameHarmonyPoints, df$red_endGameHarmonyPoints))
}

# How many alliances had 4 harmony points?
suppressWarnings(
    sum(unlist(sapply(cresc_matches, get_harmonies)) == 4, na.rm = TRUE)
)
```

# WOW Team Counts

```{r}
load("../data/teams_yearly_2024.rda")

fma_teams <- c()

# split into two loops bc 2019 was the first year of "FMA"
for (i in 2012:2018){
    fma_teams <- c(fma_teams, district_teams(paste0(i, "mar"), keys = TRUE))
}

for (i in 2019:2024){
    fma_teams <- c(fma_teams, district_teams(paste0(i, "fma"), keys = TRUE))
}

fma_teams <- unique(fma_teams)

wow <- teams_yearly %>%
    filter(year >= 2012) %>%
    filter(state_prov %in% c("OH", "PA", "WV", "Ohio",
                             "Pennsylvania", "West Virginia")) %>%
    filter(!(key %in% fma_teams)) %>%
    mutate(
        state_prov = case_when(
            state_prov == "PA" ~ "Pennsylvania",
            state_prov == "WV" ~ "West Virginia",
            state_prov == "OH" ~ "Ohio",
            TRUE ~ state_prov
        )
    ) %>%
    mutate(
        state_prov = case_when(
            state_prov == "Pennsylvania" ~ "W. Pennsylvania",
            TRUE ~ state_prov
        )
    ) %>%
    group_by(year, state_prov) %>%
    summarize(count = n())

ggplot(wow, aes(x = as.character(year), y = count, fill = state_prov)) +
    geom_bar(stat = "identity", position = position_dodge(), color = "black") +
    labs(title = "WOW Annual Team Count",
         subtitle = "2012 - 2024",
         x = "Year", y = "Count",
         fill = "State",
         caption = "Data from frc-events API via TheBlueAlliance") +
    theme_bw()

```

# RAS Sustainability

teams who won RAS
- rookie year
- first Impact win

Questions for Zaque: 
- How did you account for "Rookie All-Star" as opposed to "Rookie All Star"
- Did you count awards like "Chairman's Award Finalist"? Did you count "Chairmans Award"?

```{r}
all_team_awards <- readRDS("../data/awards/all_team_awards.rds")

extract_cols <- function(team){
    if (nrow(team) == 0) return(
        data.frame(
            ras_year = NA, first_impact = NA, n_impacts = 0, first_ei = NA,
            n_eis = 0
        )
    )
    return(data.frame(
        ras_year = team$year[grep("Rookie All[- ]Star", team$name)[1]], 
        first_impact = team$year[grep("Chairman'?s|Impact", team$name)[1]],
        n_impacts = length(grep("Chairman'?s|Impact", team$name)),
        first_ei = team$year[grep("Engineering Inspiration", team$name)[1]],
        n_eis = length(grep("Engineering Inspiration", team$name))
    ))
}

df <- lapply(all_team_awards, extract_cols)
tmp <- names(df)
df <- Reduce(rbind, df)
df$id <- tmp

ras <- df[!is.na(df$ras_year), ]
ras$yrs_to_impact <- ras$first_impact - ras$ras_year
ras$yrs_to_ei <- ras$first_ei - ras$ras_year

hist(ras$yrs_to_impact, 
     main = "Years between RAS and Impact", xlab = "Years", ylab = "# of Teams")
hist(ras$yrs_to_ei, 
     main = "Years between RAS and EI", xlab = "Years", ylab = "# of Teams")
```

