---
title: "WLS Bin Hyperparameter Tuning"
format: html
---

Applying hyperparameter tuning to optimze the number of bins for WLS methods established in [this paper](https://www.overleaf.com/read/dkmzrpspkhzb#0bd526).

```{r setup, message=FALSE, warning=FALSE}
library(devtools)
load_all()
rm(list = ls())
load("../../data/district_quals_09_24.rda")

# normalize match scores so we can do cross-week and cross-season comparisons
# this simplifies a lot of later code
matches <- lapply(
    matches, 
    function(match_df){
        scaled <- scale(c(match_df$red_score, match_df$blue_score))
        match_df$red_score <- scaled[1:nrow(match_df)]
        match_df$blue_score <- scaled[(nrow(match_df) + 1):length(scaled)]
        return(match_df)
    })

library(extrafont)
library(gt)
gos_blue <- "#337DFC"
gos_red <- "#F7041A"

gos_theme <- theme_bw() + 
    theme(text = element_text(family = "Futura"))

# fits WLS on the match dataframes
weighted_fit <- function(match_df, w){
    return(
        fit_lineup_lm(
            match_df, 
            responses = list(
                red = match_df$red_score, blue = match_df$blue_score
            ), 
            w = w
        )
    )
}

# fits the standard opr model on a regular match dataframe from TheBlueAlliance
fit_opr <- function(match_df){
    return(
        fit_lineup_lm(
            match_df, responses = list(
                red = match_df$red_score, blue = match_df$blue_score)
        )
    )
}

# this uses the "shortcut" formula for LOOCV
# see formula 5.2 in ISLR by James, Witten, Hastie, and Tibshirani (v2), pg 202
LOOCV <- function(fit){
    mean(((fit$model$response - predict(fit))/(1 - boot::glm.diag(fit)$h))^2)
}
```

The following code chunk generates the key data - the model residuals and the normalized match number.

```{r data-generation}
normalize_match_number <- function(opr_residual){
    # we know we can divide the length by 2 because there are 2 alliances/match
    alliance_corrected <- ((1:length(opr_residual) - 1) %% 
                               (length(opr_residual) / 2)) + 1
    return((alliance_corrected / length(opr_residual)) * 2)
}

fits <- lapply(matches, fit_opr)

residuals <- lapply(fits, function(fit){return(residuals(fit))})
# match percentile is the match number divided by the number of matches
match_percentile <- unlist(lapply(residuals, normalize_match_number), 
                           use.names = FALSE)
```

A reasonable grid to optimize over is simply 2 through 12. 2 is as low as you could possibly go (only 1 bin is essentially unweighted), while 12 is one bin/round, meaning many teams will not appear in bins.

```{r}
# assumes that residuals have been standardized
generate_weights <- function(n_bins, residuals, match_percentile){
    cuts <- cut(match_percentile, breaks = n_bins)
    bin_vars <- rep(0, length(unique(cuts)))
    for (i in 1:length(unique(cuts))){
        bin_vars[i] <- var(unlist(residuals)[cuts == levels(cuts)[i]])
    }
    return(1 / bin_vars)
}

mean_loocvs <- c()
n_bins <- 2:15
for (k in n_bins){
    bin_weight_fits <- lapply(
        matches, weighted_fit, 
        w = generate_weights(k, residuals, match_percentile)
    )
    mean_loocvs <- c(mean_loocvs, mean(sapply(bin_weight_fits, LOOCV)))
}
names(mean_loocvs) <- paste0("Bins = ", n_bins)

viz <- data.frame(bins = n_bins, mean_loocvs = mean_loocvs)

ggplot(viz, aes(x = n_bins, y = mean_loocvs)) + 
    geom_point(col = gos_red) + 
    labs(title = "Hyperparameter Tuning for Bins", 
         x = "Bin #", y = "LOOCV Mean") + 
    scale_x_continuous(breaks = n_bins) + 
    gos_theme

cat("The minimum occurs at", names(which.min(mean_loocvs)), 
    "with LOOCV:", mean_loocvs[which.min(mean_loocvs)], "\n")
```

**Takeaway**: Really any number between 6 and 12 will be fine, but 12 does technically come out on top.
